Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-06 WSSSPE3/report/WSSSPE3_review.tex
edits to Hetherington LT

\item \textbf{James Hetherington, Jonathan Cooper, Robert Haines, Simon
Hettrick, James Spencer, Mark Stillwell, Mike Croucher, Christopher Woods and
Susheel Varma: \textit{An update from UK Research Software Engineers}}
James Hetherington started by listing some of the problems research software
faces, which include low levels of reuse and poor standard of verification. For
a long time, technical solutions to such technical problems were focus of the
e-Research community, and include research software distributions, grids,
middleware and workflows. Some limited adoption can be seen today in research
communities, but the main problems have not been solved to a sufficient level.
James Hetherington hypothesises that instead of technical solutions, social
innovation is needed: a new role in the academic system focused on research
software that combines the best of a craftsperson and a scholar. However,
social innovation in centuries-old institutions is hard. Alternatives to such a
new role would have to include rewards for good research software, recognition
of software as academic output, and rejecting submissions based on
unreproducible computational results. Advantages of research software
engineering (RSE) groups include the possibility of training in reproducible
computational research, providing collaborations for for researchers who don't
want to be programmers, and synergies with research computing platforms. The
success of such a group would be measured by the output and quality of the
research software, and could be part of Research Computing or faculty. Members
would not be independent researchers, but would have to have a research
background. An attempt to form a community of RSE groups within the UK has been
under way for several years, including funding from the UK Research Council. An
open question is whether this approach can be adopted also in other countries,
including the USA.

>>>>

\item \textbf{James Hetherington, Jonathan Cooper, Robert Haines, Simon
Hettrick, James Spencer, Mark Stillwell, Mike Croucher, Christopher Woods and
Susheel Varma: \textit{An update from UK Research Software Engineers}}
James Hetherington started by listing some of the problems research software
faces, which include low levels of reuse and poor standard of verification. For
a long time, technical solutions to such technical problems were focus of the
eResearch community, including research software distributions, grids,
middleware and workflows. Some limited adoption can be seen today in research
communities, but the main problems have not been solved to a sufficient level.
James Hetherington hypothesizes that instead of technical solutions, social
innovation is needed: a new role in the academic system focused on research
software that combines the best parts of a craftsperson and a scholar. However,
social innovation in centuries-old institutions is hard. Alternatives to such a
new role would have to include rewards for good research software, recognition
of software as academic output, and rejecting submissions based on
unreproducible computational results. Some advantages of research software
engineering (RSE) groups include the possibility of training in reproducible
computational research, providing collaborations for for researchers who don't
want to be programmers, and synergies with research computing platforms. The
success of such a group would be measured by the output and quality of the
research software, and could be part of Research Computing or faculty. Members
would not be independent researchers, but would have to have a research
background. An attempt to form a community of RSE groups within the UK has been
underway for several years, including funding from the UK Research Council. An
open question is whether this approach can be adopted also in other countries,
including the USA.
